Article Title

### 2013-12-16 plos.tex
unfinished edits of the Discussion section.

However, neural correlations do not submit to ready or unambiguous mechanistic interpretation.  Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including  direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.

@@

Given a covariance matrix estimate $C$ (not necessarily sample covariance $C_{\sf 0}$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
\begin{equation}
    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2}
\end{equation}
Similarly, the matrix of partial correlations $P$ is computed as the negative normalized version of the \emph{precision matrix} (inverse covariance matrix) $C^{-1}$:
\begin{equation}
    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
\end{equation}
Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix.

@@

The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations are the result of direct linear interactions between pairs of observed cells and that such interactions occur only between a fraction of such pairs (Fig.~\ref{fig:02}\,C).  This assumption is enforced by reducing to zero the majority of pairwise \emph{partial correlations} in the recorded population.  While usual correlations are calculated from the marginal distribution without conditioning on all the other neurons, the partial correlation expresses the pairwise correlation conditioned on the activity of all the other cells.  If cells only exert linear effects on each other and all neurons are recorded, partial correlations express the direct interactions between neurons.  Therefore, this estimator is biased toward the assumption that correlations arise due to interactions between a subset of pairs of recorded neurons (Fig.~\ref{fig:03}C). When the partial correlation between a pair of neurons is zero, then the corresponding element of the inverse of the covariance matrix (often referred to as the \emph{precision matrix} or \emph{concentration matrix}) must be zero as well. Then the estimator is 
\begin{equation}
C_{\sf sparse} = S^{-1}
\end{equation}
where $S$ is a sparse matrix with a large fraction of zeros in its off-diagonal elements. The estimate has one hyperparameter to regulate the sparsity (fraction of off-diagonal zeros) in $S$.

Finally, we consider the fourth estimator $C_{\sf sparse+latent}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D). This estimator has the form \cite{Chandrasekaran:2010,Ma:2013}
\begin{equation}
C_{\sf sparse+latent} = (S - LL^\T)^{-1}
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a $d\times p$ matrix of factor loadings. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.

@@

In this study, we adopted \emph{negative normal log-likelihood loss}\footnote{
The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_{\sf 0})$: 
\begin{equation*}
    L(\Sigma|C_{\sf 0}) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_{\sf 0}}
\end{equation*}
}
defined as
\begin{equation}\label{eq:loss}
    \loss{C,\Sigma} = \frac 1 p\left[\ln \det C + \Tr(C^{-1}\Sigma)\right]
\end{equation}

@@

\section*{Discussion}
\subsection*{Functional connectivity from big neural data}
Transformational breakthroughs in neuroscience are anticipated to emerge from technological developments that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far.

\subsection*{Meaning of interactions}
\subsection*{Fitting probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
\subsection*{Future directions}

>>>>

However, neural correlations do not submit to ready or unambiguous mechanistic interpretation.  Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including  direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967b,Moore:1970,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.

@@

Given a covariance matrix estimate $C$ (not necessarily the sample covariance $C_{\sf 0}$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
\begin{equation}
    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2}
\end{equation}
Similarly, the matrix of partial correlations $P$ is computed as the negative normalized version of the \emph{precision matrix} (inverse covariance matrix) $C^{-1}$:
\begin{equation}
    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
\end{equation}
Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. Clearly, off-diagonal zeros in the precision matrix correspond to zero partial correlations.

@@

The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations are the result of direct linear interactions between pairs of observed cells and that such interactions occur only between a fraction of such pairs (Fig.~\ref{fig:02}\,C).  This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.  Therefore, $C_{\sf sparse}$ is biased toward the assumption that correlations arise due to linear interactions between pairs of recorded neurons (Fig.~\ref{fig:03}C). Then the estimator is 
\begin{equation}
C_{\sf sparse} = S^{-1}
\end{equation}
where $S$ is a sparse matrix with a large fraction of zeros in its off-diagonal elements. The estimate has one hyperparameter to regulate the sparsity (fraction of off-diagonal zeros) in $S$.

Finally, we consider the fourth estimator $C_{\sf sparse+latent}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D). This estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
\begin{equation}
C_{\sf sparse+latent} = (S - LL^\T)^{-1}
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a $d\times p$ matrix of factor loadings. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.

@@

In this study, we adopted \emph{negative normal log-likelihood loss}\footnote{
The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_{\sf 0})$ as
\begin{equation*}
    L(\Sigma|C_{\sf 0}) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_{\sf 0}}
\end{equation*}
}
defined as
\begin{equation}\label{eq:loss}
    \loss{C,\Sigma} = \frac 1 p\left[\ln \det C + \Tr(C^{-1}\Sigma)\right]
\end{equation}

@@

\section*{Discussion}
\subsection*{Functional connectivity from big neural data}
Transformational breakthroughs in neuroscience are anticipated to follow technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far. It far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques. Most common analyses of the functional connectivity rely on analyzing pairwise correlation coefficients, t-tests, ANCOVA, principal component analysis of correlated activity, and cross-correlations.  Traditionally, functional connectivity between regions has been defined as the level of correlation or synchronization between regions or cells. Without targeted stimulation and intracellular signals, synaptic connectivity could be inferred with some degree of confidence from the presence of sharp peaks or troughs in the cross-correlograms at offsets corresponding to typical synaptic transmission delays. However, even such effects can be explained by multiple effects besides monosynaptic connections while some strong synaptic connections do not necessarily produce a clear spike in the cross-correlogram. The situations is a bit more complicated in recording modalities, such as calcium imaging, in which only firing rates at lower temporal resolutions can be inferred and cross-correlograms with millisecond precision cannot be obtained. Therefore, some researchers have resorted to using simple correlations as proxies of functional connectivity. Yet, correlations are the end product of multiple cell interactions and, which do not always correspond to

\subsection*{Comparison with other uses of neural correlations}
\subsection*{Meaning of interactions}
\subsection*{Fitting probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
\subsection*{Future directions}
